Between Subgraph Isomorphism and Maximum Common Subgraph

### 2016-09-13 paper/paper.tex
Remove some hackery
Make figure smaller
Be a bit more abstract
More introductioning

\usepackage{placeins}

\usepackage{showframe}

@@

\begin{abstract}
    When a small pattern graph does not occur inside a larger target graph, we can ask how to find
    ``as much of the pattern as possible'' inside the target graph. In general, this is known as the maximum
    common subgraph problem, which is much more computationally challenging in practice than
    subgraph isomorphism. We introduce a restricted alternative, where we ask if all but $k$
    vertices from the pattern can be found in the target graph. This allows for the development of
    slightly weakened forms of certain invariants from subgraph isomorphism which are based upon degree and
    number of paths.  We show that when $k$ is small, weakening the invariants still retains much of
    their effectiveness. We are then able to solve this problem on the standard SIP problem
    instances used to benchmark subgraph isomorphism algorithms, despite these instances being
    too large for current maximum common subgraph algorithms to handle.
\end{abstract}
\todo{try code on MCS instances}
\section{Introduction}
\todo{talk about subgraph isomorphism applications, particularly vision}
\todo{talk about reformulating}
The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs
\citep{DBLP:conf/cp/McCreeshNPS16}. This is largely because strong inference, based upon the degrees
of vertices \citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted,
whilst filtering during search is vastly stronger.

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then
weakened forms of degree and path based filterings are still effective in pruning the initial search
space and in providing additional constraints respectively. We then show that combining these
techniques leads to a practical algorithm which can scale to work with the families of graphs
commonly used to benchmark subgraph isomorphism algorithms: depending upon the benchmark family, in
a substantial portion of ??; in ?? more cases, we can at least obtain an upper bound. This is a
significant improvement over state of the art maximum common subgraph solvers, which cannot even fit
many of these instances in RAM.
\begin{figure}
    \begin{center}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm,node distance=10mm,fill=black}]
            \node[fill=uofgcobalt] (a) {};
            \node (a1) [below left of = a] {};
            \node (b1) [below right of = a] {};
            \node (aa) [below right of = a1] {};

@@

\begin{figure}
    \begin{center}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
            \node[fill=black] (p) {};
            \node[fill=black] (b) [above of = p,node distance=10mm] {};
            \node[fill=black] (a) [left of = b] {};
            \node[fill=black] (c) [right of = b] {};
            \node[fill=black] (d) [above of = b,node distance=10mm] {};

@@

            \path (p) edge (a) edge (b) edge (c)
                (a) edge (d)
                (b) edge (c) edge (d)
                (c) edge (d);
        \end{tikzpicture}

        \vspace{5mm}

        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
            \node[fill=gray!50] (a1) {};
            \node[fill=black] (b1) [right of = a1] {};
            \node[fill=gray!50] (c1) [right of = b1] {};
            \node[fill=black] (a2) [below of = a1] {};
            \node[fill=black] (b2) [right of = a2] {};
            \node[fill=black] (c2) [right of = b2] {};
            \node[fill=gray!50] (a3) [below of = a2] {};
            \node[fill=black] (b3) [right of = a3] {};
            \node[fill=gray!50] (c3) [right of = b3] {};

@@

            \path[color=gray!50] (a1) edge (b1) edge (a2)
                (b1) edge (c1)
                (c1) edge (c2)
                (a2) edge (b2) edge (a3)
                (c2) edge (c3)
                (a3) edge (b3)
                (b3) edge (c3);
            \path[line width=0.4mm] (b1) edge (a2) edge (b2) edge (c2)
                (a2) edge (b3)
                (b2) edge (c2) edge (b3)
                (c2) edge (b3);
        \end{tikzpicture}
        \hspace{1cm}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
            \node[fill=black] (p) {};
            \node[draw=none] (b) [above of = p,node distance=10mm] {};
            \node[fill=black] (a) [left of = b] {};
            \node[fill=black] (c) [right of = b] {};
            \node[fill=black] (d) [above of = b,node distance=10mm] {};

@@

            \path (p) edge (a) edge (c)
                (a) edge (d)
                (c) edge (d);
        \end{tikzpicture}
        \caption{$P\rightarrowtail T_{1}$, $P\lessnonind{1} T_{2}$}
        \label{fig:klessexample}
    \end{center}
\end{figure}

@@

\FloatBarrier

>>>>

% \usepackage{showframe}

@@

\begin{abstract}
    When a small pattern graph does not occur inside a larger target graph, we can ask how to find
    ``as much of the pattern as possible'' inside the target graph. In general, this is known as the maximum
    common subgraph problem, which is much more computationally challenging in practice than
    subgraph isomorphism. We introduce a restricted alternative, where we ask if all but $k$
    vertices from the pattern can be found in the target graph. This allows for the development of
    slightly weakened forms of certain invariants from subgraph isomorphism which are based upon degree and
    number of paths.  We show that when $k$ is small, weakening the invariants still retains much of
    their effectiveness. We are then able to solve this problem on the standard SIP problem
    instances used to benchmark subgraph isomorphism algorithms, despite these instances being
    too large for current maximum common subgraph algorithms to handle. Finally, by iteratively
    increasing $k$, we obtain an algorithm which is also competitive for the maximum common subgraph
    problem.
\end{abstract}

\section{Introduction}
\todo{talk about subgraph isomorphism applications, particularly vision}
\todo{talk about reformulating}
The subgraph isomorphism problem is to find a copy of a small \emph{pattern} graph inside a larger
\emph{target} graph. It comes in two forms: in the non-induced variant, edges must be mapped to
edges, but the target may have ``extra edges'', whilst in the induced variant, non-edges may only be
mapped to non-edges. When a pattern cannot be found, we may wish to be given a result which maps as
many vertices of the pattern into the target as possible. In the induced case, this is known as the
maximum common induced subgraph problem (we discuss the non-induced case below). However, although
recent subgraph isomorphism algorithms are comfortable working with graphs with thousands of
vertices, the state of the art for the maximum common subgraph problem becomes computationally
infeasible at only 35 vertices when working with unlabelled graphs
\citep{DBLP:conf/cp/McCreeshNPS16}. This is largely because strong inference, based upon the degrees
of vertices \citep{DBLP:journals/ai/Solnon10} and the distances or paths between them
\citep{DBLP:conf/cp/AudemardLMGP14,DBLP:conf/cp/McCreeshP15}, is possible with subgraph isomorphism,
but not maximum common subgraph, and so the state space in the former is much more restricted,
whilst filtering during search is vastly stronger.

In this work we discuss an intermediate problem, where we must map all but $k$ vertices of the
pattern graph into the target. We show that if $k$ is reasonably small (say, between 1 and 5), then
weakened forms of degree and path based filterings are still effective in pruning the initial search
space and in providing additional constraints respectively. We then show that combining these
techniques leads to a practical algorithm which can scale to work with the families of graphs
commonly used to benchmark subgraph isomorphism algorithms: depending upon the benchmark family, we
can close a substantial portion of the instances, and in many more cases, we can at least obtain an
upper bound. This is a significant improvement over state of the art maximum common subgraph
solvers, which cannot even fit many of these instances in RAM. Finally, we show that iteratively
increasing $k$ gives a competitive algorithm for the maximum common induced subgraph problem.

\begin{figure}
    \begin{center}
        \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm,node distance=10mm,fill=black}]
            \node[fill=uofgcobalt] (a) {};
            \node (a1) [below left of = a] {};
            \node (b1) [below right of = a] {};
            \node (aa) [below right of = a1] {};

@@

\begin{figure}
    \centering
    \begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=black] (p) {};
        \node[fill=black] (b) [above of = p,node distance=10mm] {};
        \node[fill=black] (a) [left of = b] {};
        \node[fill=black] (c) [right of = b] {};
        \node[fill=black] (d) [above of = b,node distance=10mm] {};

@@

        \path (p) edge (a) edge (b) edge (c)
            (a) edge (d)
            (b) edge (c) edge (d)
            (c) edge (d);
        \end{tikzpicture}\hspace{0.5cm}\begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=gray!50] (a1) {};
        \node[fill=black] (b1) [right of = a1] {};
        \node[fill=gray!50] (c1) [right of = b1] {};
        \node[fill=black] (a2) [below of = a1] {};
        \node[fill=black] (b2) [right of = a2] {};
        \node[fill=black] (c2) [right of = b2] {};
        \node[fill=gray!50] (a3) [below of = a2] {};
        \node[fill=black] (b3) [right of = a3] {};
        \node[fill=gray!50] (c3) [right of = b3] {};

@@

        \path[color=gray!50] (a1) edge (b1) edge (a2)
            (b1) edge (c1)
            (c1) edge (c2)
            (a2) edge (b2) edge (a3)
            (c2) edge (c3)
            (a3) edge (b3)
            (b3) edge (c3);
        \path[line width=0.4mm] (b1) edge (a2) edge (b2) edge (c2)
            (a2) edge (b3)
            (b2) edge (c2) edge (b3)
            (c2) edge (b3);
    \end{tikzpicture}\hspace{0.5cm}\begin{tikzpicture}[every node/.style={circle, inner sep=0.5mm}]
        \node[fill=black] (p) {};
        \node[draw=none] (b) [above of = p,node distance=10mm] {};
        \node[fill=black] (a) [left of = b] {};
        \node[fill=black] (c) [right of = b] {};
        \node[fill=black] (d) [above of = b,node distance=10mm] {};

@@

        \path (p) edge (a) edge (c)
            (a) edge (d)
            (c) edge (d);
    \end{tikzpicture}
    \caption{$P\rightarrowtail T_{1}$, $P\lessnonind{1} T_{2}$}
    \label{fig:klessexample}
\end{figure}

@@
